---
title: "Kelp meta-analysis code"
author: "Miranda Roethler"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r include=FALSE}
#r setup
knitr::opts_chunk$set(echo = TRUE)
```


## Set up workspace: load files, libraries (`metafor`, `tidyverse`)
For an overview of what `metafor` can do, see: `vignette("diagram")`; also see Viechbauer et al. 2010 and Koricheva et al. 2013
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/Documents/Github/Kelp-meta-analysis-project") #set working directory
```
```{r echo = FALSE}
library(metafor) #load required packages
library(tidyverse)
library(writexl)

data <- read.csv("Kelp dataset.csv") #load dataset
```

Notes on data frame: 
Columns with `.t` stand for treatment, and columns with `.c` stand for control.

Ecoregions were coded in using in the Krumhansl et al. 2016 paper and courtesy of their Github (https://github.com/kelpecosystems/global_kelp_time_series/blob/master/02_sas_and_r_data_derivation_scripts/addRegionalData.R).

For this version of the data frame, effect size has already been calculated. The code to calculate effect size is below.
Data.c and Data.t represent columns in the original database that contained the raw study data.
```{effect size}
#calculate effect size
efsize <- escalc(measure = "SMD",
               n1i = data7$Sample_size.c, #group sizes or row totals
               n2i = data7$Sample_size.t, 
               m1i = data7$Data.c, #means
               m2i = data7$Data.t, 
               sd1i = data7$StDev.c + 0.00001, #standard deviations, plus 0.0001 to omit problems with absolute 0s
               sd2i = data7$StDev.t + 0.00001, 
               add=0.001, to="if0all") #add a small value to 0's so they're not 0's anymore

#add effect size calculations into the main dataset 
data8 <- bind_cols(data7, efsize)
```


## Optional: Delete data from studies that collected at multiple timepoints
Take this step unless looking at experimental duration as a variable. 
In the data file, for studies that took measurements at multiple timepoints, in the column "Timepoint", "F" means that value was taken at the end of the experiment ("final value") and "NF" means that the value was not taken at the end of the experiment ("not final value"). 
Here we will filter out all rows with "NF" values. 
```{r}
#NF timepoint
data2 <- filter(data, Timepoint == "F")
```
OR:
```{r}
data2 <- data
```

## Optional: Delete 'redundant' metrics (i.e. if a study included multiple ways to measure growth, e.g. change in weight AND change in length, just include one). This script will take the most common metric used in the dataset and use that one. The exception is for biochemical composiiton, where we have C:N ratio ranked as the highest priority metric even though it is not the most common.
```{r}
#delete redundant metrics
#make a list of all of the categories
category_list <- unique(data2$Response_category)
#take out biochemical composition because that one is going to be manually organized 
category_list <- category_list[category_list != "Biochemical composition"]
#create an empty list to store each category's response in
filteredlist <- vector(mode = "list", length = length(category_list) + 1)

for (x in 1:length(category_list)) {
  
  growth <- filter(data2, Response_category == category_list[x]) #filter out only one  category at a time
  growthlist <- growth %>% #make a list of all the variables this contains
    count(Response_variable) %>%
    arrange(desc(n)) #arrange in order from most to least entries
  
  growthcount <- growth %>% #record the number of ways used to measure this category within each study
    count(Study, Lifestage, Response_variable) %>%
    arrange(Study) %>% #arrange in order from most to least entries
    count(Study, Lifestage)
  
  growthstudy <- growth %>% 
    group_split(Study, Lifestage)
  
  output <- vector(mode = "list", length = length(growthstudy))
  
  for ( i in 1:nrow(growthcount)) {
    
    if (growthcount[i,2] > 1) {
      respvar <- unique(growthstudy[[i]]$Response_variable) %>%   
        str_c(collapse = "|") #create a character string with all of the response variables for that specific study
      
      finalvar <- str_extract(growthlist[,1], respvar) %>%
        na.omit() #figure out the order of those variables from the reference list, i.e. who appears most often in the whole dataset
      
      output[[i]] <- filter(growthstudy[[i]], Response_variable == finalvar[1])
      
      }
    
    else {
      output[[i]] <- growthstudy[[i]]
    
      }
    }
  filteredlist[[x]] <- bind_rows(output)
}

#Now do biochemical composition separately
y <- length(category_list) + 1 #this is the position Biochemical Composition should be in
biochem <- filter(data2, Response_category == "Biochemical composition") #filter out only biohcemical composition category
biochemlist <- biochem %>% #make a list of all the variables this contains
  count(Response_variable) %>%
  arrange(desc(n)) %>% #arrange in order from most to least entries
  subset(Response_variable != "C:N ratio")   #take out C:N ratio
CN <- c("C:N ratio", "NA")
biochemlist <- rbind(CN, biochemlist)#add in C:N ratio to the top
  
biochemcount <- biochem %>% #record the number of ways used to measure growth within each study
    count(Study, Lifestage, Response_variable) %>%
    arrange(Study) %>%
    count(Study, Lifestage)
  
biochemstudy <- biochem %>% 
    group_split(Study, Lifestage)
  
output <- vector(mode = "list", length = length(biochemstudy))
  
for ( i in 1:nrow(biochemcount)) {
    
    if (biochemcount[i,2] > 1) {
      respvar <- unique(biochemstudy[[i]]$Response_variable) %>%   
        str_c(collapse = "|") #create a character string with all of the response variables for that specific study
      
      finalvar <- str_extract(biochemlist[,1], respvar) %>%
        na.omit() #figure out the order of those variables from the reference list, i.e. who appears most often in the whole dataset
      
      output[[i]] <- filter(biochemstudy[[i]], Response_variable == finalvar[1])
      
      }
    
    else {
      output[[i]] <- biochemstudy[[i]]
    
      }
    }
filteredlist[[y]] <- bind_rows(output)

#at the very end:

data3 <- bind_rows(filteredlist)


```
OR:
```{r}
data3 <- data2
```



Make sure R is reading all of the columns like it should 
```{r}
#column type QAQC
#make sure R is reading all of the factors as factors and not characters
data3$Study <- as.factor(data3$Study)
data3$Family <- as.factor(data3$Family)
data3$Genus <- as.factor(data3$Genus)
data3$Species <- as.factor(data3$Species)
data3$Stressor <- as.factor(data3$Stressor)
data3$Lifestage <- as.factor(data3$Lifestage)
data3$Response_category <- as.factor(data3$Response_category)
data3$pCO2_uatm.t <- as.numeric(data3$pCO2_uatm.t)
data3$pCO2_uatm.c <- as.numeric(data3$pCO2_uatm.c)
data3$Response_variable <- as.factor(data3$Response_variable)
data3$ECOREGION <- as.factor(data3$ECOREGION)
data3$PROVINCE <- as.factor(data3$PROVINCE)
data3$REALM <- as.factor(data3$REALM)
```



#test for publication bias
```{r publication}
###### publication bias ###########

## 
#funnel plot for all OA metrics to test for publication bias
rma_oa <-rma(yi,vi, method="REML", data=subset(data3, Stressor == 'OA'))
funnel(rma_oa, pch = 20) 
#test for funnel asymmetry
regtest(rma_oa, model = "lm") 

#funnel plot for all T metrics to test for publication bias
rma_t <-rma(yi,vi, method="REML", data=subset(data3, Stressor == 'T'))
funnel(rma_t, pch = 20) 
#test for funnel asymmetry
regtest(rma_t, model = "lm") 

#funnel plot for all OA + T metrics to test for publication bias
rma_oat <-rma(yi,vi, method="REML", data=subset(data3, Stressor == 'OA + T'))
funnel(rma_oat, pch = 20) 
#test for funnel asymmetry
regtest(rma_oat, model = "lm") 

```


#continuous variables: pco2, temperature
```{r modeling}
#pCO2
rma.pCO2 <- rma.mv(yi = yi, V = vi, data = data3, 
                    random = list( ~ Study | Experiment,
                                  ~ 1 | Species),
                    mods = ~ deltapCO2_z + Duration_days_z, 
                    subset = ((Stressor == "OA" | Stressor == "OA + T") 
                              &  deltapCO2 != "NA" &
                                 Duration_days != "NA"),
                    method = "REML",
                    control=list(optimizer="optim", optmethod="Nelder-Mead"), 
                    sparse = TRUE, 
                    verbose = TRUE #this will display model progress
                     )
summary(rma.pCO2) 


rma.t <- rma.mv(yi = yi, V = vi, data = data3, 
                    random = list( ~ Study | Experiment,
                                  ~ 1 | Species),
                    mods = ~ deltaT_z + Duration_days_z, 
                    subset = ((Stressor == "T" | Stressor == "OA + T") 
                              &  deltaT != "NA" &
                                 Duration_days != "NA"),
                    method = "REML",
                    control=list(optimizer="optim", optmethod="Nelder-Mead"), 
                    sparse = TRUE, 
                    verbose = TRUE #this will display model progress
                     )
summary(rma.t)


#pCO2
regplot(rma.pCO2, mod="deltapCO2", xlab="Increase in pCO2 (uatm)", 
        ylim = c(-10, 10), 
        refline = 0)


#temperature
regplot(rma.t, mod="deltaT", xlab="Increase in temperature (deg C)", 
        ylim = c(-10, 20), 
        refline = 0)


```

Duration
```{r modeling}
rma.dur.oa <-  rma.mv(yi = yi, V = vi, data = data3, 
                    random = list( ~ Study | Experiment,
                                  ~ 1 | Species),
                     subset = (Stressor == "OA" & 
                                 deltapCO2 != "NA" & 
                                 Duration_days != "NA"),
                    mods = ~ Duration_days_z + deltapCO2_z, 
                    method = "REML",
                    control=list(optimizer="optim", 
                                 optmethod="Nelder-Mead"), 
                    sparse = TRUE, 
                    verbose = TRUE 
                     )
summary(rma.dur.oa) 

#OAT
rma.dur.oat <-  rma.mv(yi = yi, V = vi, data = data3, 
                    random = list( ~ Study | Experiment,
                                  ~ 1 | Species),
                     subset = (Stressor == "OA + T" & 
                                 deltapCO2 != "NA" &
                                 deltaT != "NA" &
                                 Duration_days != "NA"),
                    mods = ~ Duration_days_z + deltaT_z + deltapCO2_z, 
                    method = "REML",
                    control=list(optimizer="optim", 
                                 optmethod="Nelder-Mead"), 
                    sparse = TRUE, 
                    verbose = TRUE 
                     )
summary(rma.dur.oat)


#T
rma.dur.t <-  rma.mv(yi = yi, V = vi, data = data3, 
                    random = list( ~ Study | Experiment,
                                  ~ 1 | Species),
                     subset = (Stressor == "T" & 
                                 deltaT != "NA" &
                                 Duration_days != "NA"),
                    mods = ~ Duration_days_z + deltaT_z, 
                    method = "REML",
                    control=list(optimizer="optim", 
                                 optmethod="Nelder-Mead"), 
                    sparse = TRUE, 
                    verbose = TRUE 
                     )
summary(rma.dur.t)

#plot it
regplot(rma.dur.t, mod="Duration_days", xlab="Experiment duration (days)", 
        ylim = c(-10, 20),
        refline = 0)


regplot(rma.dur.oa, mod="Duration_days", xlab="Experiment duration (days)", 
        ylim = c(-10, 20),
        refline = 0)


regplot(rma.dur.oat, mod="Duration_days", xlab="Experiment duration (days)", 
        ylim = c(-10, 20),
        refline = 0)

```


Example of R script for a categorical model using the life stage models as an example.
```{r modeling}
rma.life.oa <- rma.mv(yi = yi, V = vi, data = data3, 
                     random = list( ~ Study | Experiment,
                                   ~ 1 | Species),
                     subset = (Stressor == "OA" & 
                                 deltapCO2 != "NA" & 
                                 Duration_days != "NA"
                               ),
                     mods = ~ Lifestage + Duration_days_z + 
                       deltapCO2_z - 1, 
                     method = "REML",
                     control=list(optimizer="optim", optmethod="Nelder-Mead"),
                     verbose = TRUE, 
                     sparse = TRUE) 

summary(rma.life.oa)


rma.life.t <- rma.mv(yi = yi, V = vi, data = data3, 
                     random = list( ~ Study | Experiment,
                                   ~ 1 | Species),
                     subset = (Stressor == "T" & 
                                 deltaT != "NA" & 
                                 Duration_days != "NA"),
                     mods = ~ Lifestage + Duration_days_z + 
                       deltaT_z - 1, 
                     method = "REML",
                     control=list(optimizer="optim", optmethod="Nelder-Mead"),
                     verbose = TRUE, 
                     sparse = TRUE) 

summary(rma.life.t)


rma.life.oat <- rma.mv(yi = yi, V = vi, data = data3, 
                     random = list( ~ Study | Experiment,
                                   ~ 1 | Species),
                     subset = (Stressor == "OA + T" & 
                                 deltaT != "NA" & 
                                 deltapCO2 != "NA" &
                                 Duration_days != "NA"),
                     mods = ~ Lifestage + Duration_days_z + 
                       deltaT_z + deltapCO2_z - 1, 
                     method = "REML",
                     control=list(optimizer="optim", optmethod="Nelder-Mead"),
                     verbose = TRUE,
                     sparse = TRUE) 

summary(rma.life.oat)
```

